Can Cultural Transmission Explain The Evolution of Cooperative Behavior?

### 2020-03-14 ms/main.tex

With only vertical transmission eq.~\ref{eq:vertical_oblique_transmission} becomes
$\hat{p} =  \tilde{p}$,
and eq.~\ref{eq:nextgen_parents} for the change in frequency $p'$ of phenotype $A$ among parents can be written as

\begin{equation} 
\begin{split}\label{eq:nextgen_parents_vertical_only}
\bar{w} \tilde{p}' 
& = p^2 (1+b-c) (1 - (1-p) (1-\alpha) T_B) \\
& + p(1-p) (1-c) (p (1-\alpha) T_B + 1 - T_B) \\
& + p(1-p) (1+b) (p (1-\alpha) + \alpha) T_A \\
& + (1-p)^2 p (1-\alpha) T_A
\end{split}
\end{equation}

\textbf{Result 1: Vertical and horizontal transmission of cooperation.}
If 
\begin{equation}\label{result1}
c \cdot (1-T_B) < b \cdot  \alpha T_A  + (T_A - T_B) [1 + b (p(1-\alpha) - 1)]
\end{equation}
then $\tilde{p}' > \tilde{p}$, and the frequency of phenotype $A$ among parents increases.

@@

\underline{Special case 2 - $\alpha =1$:} in this case, horizontal transmission can only occur as a result of interaction.
\\Eq. \eqref{result0} becomes:
\begin{equation} 
\begin{split} \label{result3}
c(1-T_B)  < & \, T_A -T_B +b(T_A-T_B)
\end{split}
\end{equation}
Therefore, cooperative behavior will take over the population when eq. \eqref{result3} is satisfied. 
This result is in accordance with the result presented by Lewin-Epstein at el. \citep{Ohad}. 
\\\underline{Special case 3 - $\alpha = 0$:} in this case horizontal transmission can not occur as a result of interaction.  
Then, eq. \eqref{result0} becomes:
\begin{equation} 
\begin{split} \label{result4}
0 < & \,  - bq(1-T_B)- bT_B-c(1-T_B)-T_B
\end {split}
\end{equation}
Which is always false because all of the elements are negative. Therefore, cooperation will eventually disappear from population.
\\Those are the results we have so far. However, there is a still work need to be done. \\
We will analyze the equilibrium points in this model. We know that $p=1$ is an equilibrium point because if the all individuals in the population are cooperators then in next generation all individual will be cooperator as well. Same for $p=0$. We will check if there are any more equilibrium points. For each equilibrium point we will determine whether it is stable equilibrium or not. 
\\So far, we assumed that $v=1$. We believe it will be interesting to test the effect of oblique transmission ($v\neq 1$) since oblique transmission is common in cultural evolution models. Therefore, we are planning to simulations to investigate the case when $v\neq 1$.

>>>>

With only vertical transmission eq.~\ref{eq:horizontal} becomes
$\hat{p} =  \tilde{p}$,
and eq.~\ref{eq:nextgen_parents} for the change in frequency $p'$ of phenotype $A$ among parents can be written as

\begin{equation} 
\begin{split}\label{eq:nextgen_parents_vertical_only}
\bar{w} \tilde{p}' 
& = \tilde{p}^2 (1+b-c) (1 - (1-\tilde{p}) (1-\alpha) T_B) \\
& + \tilde{p}(1-\tilde{p}) (1-c) (\tilde{p} (1-\alpha) T_B + 1 - T_B) \\
& + \tilde{p}(1-\tilde{p}) (1+b) (\tilde{p} (1-\alpha) + \alpha) T_A \\
& + (1-\tilde{p})^2 \tilde{p} (1-\alpha) T_A
\end{split}
\end{equation}

\textbf{Result 1: Vertical and horizontal transmission of cooperation.}
If 
\begin{equation}\label{eq:unequal_transmission}
c \cdot (1-T_B) < b \cdot  \alpha T_A  + (T_A - T_B) [1 + b (p(1-\alpha) - 1)]
\end{equation}
then $\tilde{p}' > \tilde{p}$, and the frequency of phenotype $A$ among parents increases.

@@

\underline{Special case 2 - $\alpha =1$:} in this case, horizontal transmission can only occur as a result of interaction.
\\Eq.~\ref{eq:unequal_transmission} becomes:
\begin{equation} 
\begin{split} \label{eq:horizontal_transmission_only_after_interaction}
c(1-T_B)  < & \, T_A -T_B +b(T_A-T_B)
\end{split}
\end{equation}
Therefore, cooperative behavior will take over the population when eq.~\ref{eq:horizontal_transmission_only_after_interaction} is satisfied. 
This result is in accordance with the result presented by~\citet{lewin2017microbes}. 
\\\underline{Special case 3 - $\alpha = 0$:} in this case horizontal transmission can not occur as a result of interaction.  
Then, eq.~\ref{eq:unequal_transmission} becomes:
\begin{equation} 
\begin{split} \label{eq:horizontal_transmission_from_population}
0 < & \,  - bq(1-T_B)- bT_B-c(1-T_B)-T_B
\end {split}
\end{equation}
Which is always false because all of the elements are negative. Therefore, cooperation will eventually disappear from population.
\\Those are the results we have so far. However, there is a still work need to be done. \\
We will analyze the equilibrium points in this model. We know that $p=1$ is an equilibrium point because if the all individuals in the population are cooperators then in next generation all individual will be cooperator as well. Same for $p=0$. We will check if there are any more equilibrium points. For each equilibrium point we will determine whether it is stable equilibrium or not. 
\\So far, we assumed that $v=1$. The next section discuss the effect of oblique transmission ($v\neq 1$) which is common in cultural evolution models. Start by looking on special case where $v=0$. In this case eq.~\ref{eq:horizontal} becomes $\hat{p}=p$ and eq.~\ref{eq:nextgen_parents} becomes:
\begin{equation} 
\begin{split}\label{eq:nextgen_parents_oblique_only}
\bar{w} p' 
& = p^2 (1+b-c) (1 - (1-p) (1-\alpha) T_B) \\
& + p(1-p) (1-c) (p (1-\alpha) T_B + 1 - T_B) \\
& + p(1-p) (1+b) (p (1-\alpha) + \alpha) T_A \\
& + (1-p)^2 p (1-\alpha) T_A
\end{split}
\end{equation}
The frequency of phenotype A among parents increases if $\tilde{p}'>\tilde{p}$. From eq.~\ref{eq:nextgen_parents_oblique_only} applies that $\tilde{p}'>\tilde{p}$ is satisfy when
